---
title: "Technical Appendix"
author: Hannah Ha Khuong
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
 html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---
```{r include = FALSE}
#Course: BUAN 5210
#Title: Quick Project 1  
#Author: Hannah Ha Khuong
#Date: Jan 26, 2019
```

```{r include = FALSE}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 

# Clear environmet of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)
```

```{r warning = FALSE, message = FALSE}
#Load packages 
library(dplyr)
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(reshape2)

```

#Load, transform and summarize data 
```{r}
#Load data
data <- read_csv("5210_W19_QP1_Data.csv")

#Make profit and revenue in a same column
meltdata <- melt(data,id.vars = c("Region", "Sub-Category","Category", "Segment") , measure.vars = c("Revenue", "Profit"))
```


```{r message = FALSE}
#Graph functions
boxplot <- function(b){
  b + geom_boxplot(stat = "boxplot")+
    guides(fill=FALSE)+
    theme_bw() + theme(panel.border = element_blank(), 
                       panel.grid.major = element_blank(),
                       panel.grid.minor = element_blank()) 
}
count_bargraph <- function(b){
  b + geom_bar(fill = "tan", color = "black") +
    guides(fill=FALSE)+
    theme_bw() + theme(panel.border = element_blank(), 
                       panel.grid.major = element_blank(),
                       panel.grid.minor = element_blank()) + 
    labs(y ="Count")
}

bargraph <- function(b){
  b + stat_summary(geom = "bar", fun.y = "sum", color = "black", fill = "tan") +
    guides(fill=FALSE) +
    theme_bw() + theme(panel.border = element_blank(), 
                       panel.grid.major = element_blank(),
                       panel.grid.minor = element_blank())+ 
    scale_fill_brewer(palette="Pastel1")
}

bargraph_dodge <- function(b){
  b + stat_summary(geom = "bar", fun.y = "sum", position = "dodge", color = "black") +
    theme_bw()+ theme(panel.border = element_blank())+
    scale_fill_brewer(palette="Pastel2")
}

distribution <- function(b){
  b + geom_density() +
    theme_bw()
}

scatterplot <- function(b){
  b + geom_point()  +
    theme_bw()
}
```

```{r message = FALSE}
summary(data)
```

* Most order contains 2-5 items. 
* Revenue is highly skewed to the right. 
* The range of profits is wide, but the mean and median show that the profitablity is not high (mean is $28 and median is $8)
* Discount is often at 15-20%. 

#EDA Note 
* This EDA contains 3 main parts:
    + Univariate analysis
    + Multivariate analysis
    + Questions 
  
* The values displayed on graphs for Revenue and Profit are **sums of Revenue** and **sums of Profits**. I want to understand how much total revenue/profit a region/category/subategory/segment can generate, instead of the mean or median of all orders. 

#Univariate analysis 

## Categorical variables 

###Region

```{r message = FALSE}
#Make summary table, count bargraph
Region_tab <- rbind(table(data$Region),
                    prop.table(table(data$Region))*100)
row.names(Region_tab) <- c("Count", "Percent")
round(Region_tab, 2)

count_bargraph(ggplot(data, aes(Region)))
```

* South region has the fewest orders while West region has the most orders. 

###Category
```{r message = FALSE}
#Make summary table, count bargraph
Cat_tab <- rbind(table(data$Category),
                    prop.table(table(data$Category))*100)
row.names(Cat_tab) <- c("Count","Percent")
round(Cat_tab, 2)


count_bargraph(ggplot(data, aes(Category)))
```

* Office Suplies items are the most sold. 

###Sub-Category
```{r message = FALSE}
#Make summary table, count bargraph
subCat_tab <- rbind(table(data$`Sub-Category`),
                    prop.table(table(data$`Sub-Category`))*100)
row.names(subCat_tab) <- c("Count","Percent")
round(subCat_tab, 2)


count_bargraph(ggplot(data, aes(`Sub-Category`))) + coord_flip()
```

* The numbers of machines and copiers sold are the lowest.
* The number of papers and binders sold are the highest. 

###Segment
```{r message = FALSE}
#Make summary table, count bargraph
seg_tab <- rbind(table(data$Segment),
                    prop.table(table(data$Segment))*100)
row.names(seg_tab) <- c("Count","Percent")
round(seg_tab, 2)


count_bargraph(ggplot(data, aes(Segment)))
```
 
* The number of orders from Customers is the highest, then comes the number of orders in Corporate and Home Office segment. 

### Discount
```{r message = FALSE}
#Make summary table

count_bargraph(ggplot(data, aes(Discount)))
```

* Most discount is at 20% level.
* The number of items on discount is really high.

##Continuous variables 

### Revenue
```{r message = FALSE}
#Create density graph
distribution(ggplot(data, aes(Revenue)))

```

* Revenue is highly skewed to the right.
* Most items bring revenue of less than $2,500. 

### Profit 
```{r message = FALSE}
#Create density graph
distribution(ggplot(data, aes(Profit)))
```

* Distribution of Profits is leptokurtic. Most orders' profits is in the range of -$1,000 to $1,000. 

#Multivariate Analysis 

```{r message = FALSE} 
#Creat summary function for Profit 
sum_profit <- sum(data$Profit)

summary_function <- function(var) {
  var <- enquo(var)
  print(var)
  data %>% 
  group_by(!!var) %>%
  summarize(
    count = n(),
    profit_percentage = round(((sum(Profit) / sum_profit)*100), 2),
    mean = mean(Profit),
    sum = sum(Profit),
    min = min(Profit),
    max = max(Profit)
  )
}

#Creat summary function for Revenue 
sum_revenue <- sum(data$Revenue)

rev_summary_function <- function(var) {
  var <- enquo(var)
  print(var)
  data %>% 
  group_by(!!var) %>%
  summarize(
    count = n(),
    revenue_percentage  = round(((sum(Revenue) / sum_revenue)*100), 2),
    mean = mean(Revenue),
    sum = sum(Revenue),
    min = min(Revenue),
    max = max(Revenue)
  )
}

```


##Region 
###Profit
```{r message = FALSE}
#Summary and boxplot
summary_function(Region)
boxplot(ggplot(data, aes(x = Region, y = Profit))) +labs(title = "Profit in four regions")
```

###Revenue 
```{r message = FALSE}
#Summary and boxplot
rev_summary_function(Region)
boxplot(ggplot(data, aes(x = Region, y = Revenue)))+labs(title = "Revenue in four regions")

```

###Sum of revenue and profit in relation to each other
```{r message = FALSE}
#Cluster bargraph with sum of profits and revenue 
bargraph_dodge(ggplot(meltdata, aes(x = Region, y = value, group = variable, fill = variable))) + 
  labs(y = "Monetary Value", title = "Revenue and Profit in four regions") + 
  guides(fill= guide_legend("Monetary value"))


```

* West region has the highest total profit while Central and South region have the lowest profits.
* West region also has the highest total revenue. South region has the least revenue. 
    + --> a lot of sales in Central do not generate profits, compared to South region. 

##Category

###Profit
```{r message = FALSE}
#Summary and boxplot
summary_function(Category)
boxplot(ggplot(data, aes(x = Category, y = Profit)))
```

###Revenue
```{r message = FALSE}
#Summary and boxplot
rev_summary_function(Category)
boxplot(ggplot(data, aes(x = Category, y = Revenue)))

```

###Sum of revenue and profit in relation to each other
```{r message = FALSE}
#Cluster bargraph with sum of profits and revenue 
bargraph_dodge(ggplot(meltdata, aes(x = Category, y = value, group = variable, fill = variable))) + 
  labs(y = "Monetary Value", title = "Revenue and Profit in different Categories") + 
  guides(fill= guide_legend("Monetary value"))

```

* Technology category is the most profitable while Furniture category brings back the lowest profits. 
* Furniture has higher revenue than Home Office --> Furniture is selling without profits on a lot of orders, compared to Home Office category. 

##Sub-category 

###Profit
```{r message = FALSE}
#Summary and boxplot
summary_function(`Sub-Category`)
boxplot(ggplot(data, aes(x = `Sub-Category`, y = Profit))) + 
  coord_flip()
```

###Revenue
```{r message = FALSE}
#Summary and boxplot
rev_summary_function(`Sub-Category`)
boxplot(ggplot(data, aes(x = `Sub-Category`, y = Revenue))) + coord_flip()
```

###Sum of revenue and profit in relation to each other
```{r message = FALSE}
#Cluster bargraph with sum of profits and revenue 
bargraph_dodge(ggplot(meltdata, aes(x = `Sub-Category`, y = value, group = variable, fill = variable))) + 
  labs(y = "Monetary Value", title = "Revenue and Profit in different Sub-Categories") + 
  guides(fill= guide_legend("Monetary value")) + coord_flip()


```

* There are 3 subcategories that have negative profits: Table, Supplies, and Book cases. 
* Subcategory Table is selling at the company's loss, even though this subcategory has a moderate revenue compared to other subcategories. 
* Copiers do not have high revenue, but has the highest profit --> the profits per order on this subcategory is really high.  
* Papers has the best profit/revenue ratio. 
* Accessories also has a good profit/revenue ratio. 

##Segment 

###Profit
```{r message = FALSE}
#Summary and boxplot
summary_function(Segment)

boxplot(ggplot(data, aes(x = Segment, y = Profit)))
```

###Revenue
```{r message = FALSE}
#Summary and boxplot
rev_summary_function(Segment)
 
boxplot(ggplot(data, aes(x = Segment, y = Revenue)))
```

###Sum of revenue and profit in relation to each other
```{r echo = FALSE, message = FALSE}
#Cluster bargraph with sum of profits and revenue 
bargraph_dodge(ggplot(meltdata, aes(x = Segment, y = value, group = variable, fill = variable))) + 
  labs(y = "Monetary Value", title = "Revenue and Profit in different Segment") + 
  guides(fill= guide_legend("Monetary value"))
```


*Customer segment has the highest revenue and profits. 


#Questions proposed

## Is discount making company lose money? 

### Correlational Analysis 

```{r message = FALSE}
#Scatterplot to see relationship between profit, revenue and discount
scatterplot(ggplot(data, aes(x = Profit, y = Revenue, color = Discount))) +
    stat_smooth(method = lm)
```
 
* The less revenue an order is, the less profit company can make. Revenue and Profit have a positive correlation. 
* Most of negative profit is related to high discount. 

```{r message = FALSE}
#Scatterplot to see relationship between profit, revenue and discount
c2 <- scatterplot(ggplot(data, aes(x = Discount, y = Profit)))

c1 <- scatterplot(ggplot(data, aes(x = Discount, y = Revenue)))

grid.arrange(c2, c1, 
             nrow = 1, 
             top = "Discount correlation with Profit and Revenue")
```

* Greater discount makes greater loss for the company while revenue is not necessarily higher.

### Which category/subcatetory is on discount? 

```{r message = FALSE}
#Category
data %>% 
  group_by(Category) %>%
  summarize(
    mean= mean(Discount),
    median = median(Discount),
    min = min(Discount),
    max = max(Discount) 
  )

d1 <- ggplot(data, aes(x = Category, y = Profit, fill = Discount)) + geom_col(width = 0.5) + coord_flip() +
    theme_bw() + theme(panel.border = element_blank()) + guides(fill= FALSE)

#Subcategory
data %>% 
  group_by(`Sub-Category`) %>%
  summarize(
    mean = mean(Discount),
    median = median(Discount),
    min = min(Discount),
    max = max(Discount) 
  )

d2<- ggplot(data, aes(x = `Sub-Category`, y = Profit, fill = Discount)) + geom_col() + coord_flip() +
    theme_bw() + theme(panel.border = element_blank()) +labs(x = "Sub-category")

#Arrange and save image 
grid.arrange(d1, d2, 
             nrow = 2,
             top = "Which category/subcategory on discount the most? \n Category 'Office Supplies' and Sub-category 'Binders' & 'Machines' are on sale the most.",
             heights = c(1/3, 2/3))

```

```{r message = FALSE}
#Graph average discount level for each Subcategory
ggplot(data, aes(x = `Sub-Category`, y = Discount)) + 
         stat_summary(geom = "bar", fun.y = "mean", color = "black", fill = "tan", width = 0.7) +
    guides(fill=FALSE) +
    theme_bw() + 
    theme(panel.border = element_blank(), 
                       panel.grid.major = element_blank(),
                       panel.grid.minor = element_blank()) + 
    scale_fill_brewer(palette="Pastel1")+
    coord_flip() +
    labs(title = "Average Discount level of each Sub-category") 
```


* Most subcategories has discount at less than 10% level. 
* Binders has the highest discount level at 37% on average. 
* Tables, Bookcases and Machines also have high level of discount at more than 20%. 
    + Almost half of Machines sales are on discount.
    + About 2/3 of Tables sales are on discount. 
* Appliances does not discount frequently, but they offer high discount if this subcategory goes on sale. 

## Which segment should company focus on? 
```{r message = FALSE}
#Segment and profit 
data %>% 
  group_by(Segment) %>%
  summarize(
    mean = mean(Profit),
    sum = sum(Profit),
    min = min(Profit),
    max = max(Profit) 
  )

#Draw bargraph
bargraph_dodge(ggplot(data, aes(x = Segment, y = Profit, fill = Region))) +
  labs(title ="Which segment is the most profitable?",
       subtitle = "Segment 'Consumer' is the most profitable, but profits depend on regions") +
  scale_fill_brewer(palette="Pastel1")
  
#Draw col graphs to understand which segment discount more
ggplot(data) + geom_col(aes(x = Segment, y = Profit, fill = Discount)) + coord_flip() +
    theme_bw() + theme(panel.border = element_blank()) + ggtitle("Segment Costumer yields most profit, but also\n goes on discount more frequently")
```

* Customers segment is the most profitable one, especially in the West region.
* Central region's best segment is Corporate segment. 
* South region should not focus on Home Office segment.

## Which caegory/ sub-category to invest or drop based on segment
### Category
```{r message = FALSE}
#Category and Profit, divided by regions
data %>% 
  group_by(Segment, Category) %>%
  summarize(
    mean_profit = mean(Profit),
    median_profit = median(Profit),
    min_profit = min(Profit),
    max_profit = max(Profit) 
  )

#Cluster bargraph
bargraph_dodge(ggplot(data, aes(x = Category, y = Profit, fill = Segment))) + coord_flip() +
  labs(title = "Profit by Category, divided in four regions",subtitle = "Furniture is the least profitable in Central region")
```

* The only category is losing money is Furniture sold in Center region. 

### Sub-category
```{r message = FALSE}
#Subcategory and Profit, divided by regions
data %>% 
  group_by(Segment, `Sub-Category`) %>%
  summarize(
    mean_profit = mean(Profit),
    median_profit = median(Profit),
    min_profit = min(Profit),
    max_profit = max(Profit) 
  )
#Cluster bargraph
bargraph_dodge(ggplot(data, aes(x = `Sub-Category`, y = Profit, fill = Segment))) + coord_flip()+
  labs(x = "Sub-Category", 
       title = "Profits by Sub-Categories in different Segment ") 
ggsave("subcat-profit-segment.png")


```

# Graph for memorandum 

```{r message = FALSE }
#Create names and order of graph
subcat_order <- c("Tables", "Bookcases", "Supplies", "Fasteners", "Machines", "Phones", "Copiers", "Binders")

#Copy dataframe 
select_subcat <- data

#Make graphs for memo 
select_subcat %>% 
  filter(`Sub-Category` == "Tables"|`Sub-Category` == "Supplies"|`Sub-Category` == "Fasteners"|
           `Sub-Category` =="Bookcases"|`Sub-Category` == "Phones"| `Sub-Category` =="Machines"|
           `Sub-Category` == "Copiers"|`Sub-Category` =="Binders") %>% 
  ggplot(aes(x = factor( `Sub-Category`, levels=subcat_order), y = Profit, fill = Segment))+
  stat_summary(geom = "bar", fun.y = "sum", position = "dodge", color = "black") +
    theme_bw()+ 
  theme(panel.border = element_blank(), 
                       panel.grid.major = element_blank(),
                       panel.grid.minor = element_blank())+
    scale_fill_brewer(palette="Pastel2") + coord_flip()+
  labs(x = "Sub-Category", 
       title = "Graph 1. Total Profits by Sub-Categories in different Segment ") 

ggsave("select-subcat-segment.png")



```

```{r message = FALSE}
#Make second graph for memo
select_subcat %>% 
  filter(`Sub-Category` == "Tables"|`Sub-Category` == "Supplies"|`Sub-Category` == "Fasteners"|
           `Sub-Category` =="Bookcases"|`Sub-Category` == "Phones"|`Sub-Category` =="Machines"|
           `Sub-Category` == "Copiers"|`Sub-Category` =="Binders") %>% 
  ggplot(aes(x = factor(`Sub-Category`, levels=subcat_order), y = Discount, fill = Segment))+
  stat_summary(geom = "bar", fun.y = "mean", position = "dodge", color = "black", width = 0.75) +
    theme_bw()+ 
    theme(panel.border = element_blank(), 
                       panel.grid.major = element_blank(),
                       panel.grid.minor = element_blank()) + coord_flip()  + 
  labs(x = "Sub-Category", 
       title = "Graph 2. Average Discount by Sub-Categories in different Segment") +
  scale_fill_brewer(palette="Pastel2")
ggsave("subcat-discount-segment.png")

```

